Racial and Ethnic Differences in ETI

### 2023-08-04 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
ETI combined race numbers

```{r libraries}
#| include: false
library(tidyverse)
```

```{r data cleaning}
# Most cleaning performed in GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R
load("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
```

# Table

```{r}
#| label: tbl-planets
#| tbl-cap: Planets

```

@@

```{r}
#| label: fig-planets
#| fig-cap: Planets

```

>>>>

```{r data cleaning}
#| include: false
source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
load("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
```

```{r libraries}
#| include: false
library(Hmisc)
library(tidyverse)
library(gtsummary)
```

# Race and ethnicity breakdown

```{r table 1}
#| label: tbl-1
#| tbl-cap: Race by year
annual %>%
  select(ReviewYear, race, Hispanicrace) %>%
  tbl_summary(by = ReviewYear, missing_text = "Missing")
```

```{r table 2}
#| label: tbl-2
#| tbl-cap: Race and ethnicity
demo %>%
  select(race, Hispanicrace) %>%
  tbl_summary(by = Hispanicrace) %>%
  modify_header(label = paste0("**", label(annual$Hispanicrace), "**")) %>%
  add_overall()
```

Based on the above table, I decided to combine race and ethnicity with Hispanic "overriding" other racial identities. So if a participant was reported as both Black and Hispanic, they were consider Hispanic for the following analyses. Those with missing data for Hispanic ethnicity were moved to the "Mixed/Other/Unknown Race" category.

```{r combine race and eth}
demo$race <- as.character(demo$race)
demo$race[demo$Hispanicrace == "Yes"] <- "Hispanic or Latino"
demo$race[demo$Hispanicrace == "Unknown"] <- "Hispanic or Latino"
demo$race <- factor(demo$race,
  levels = c(
    "White", "Black or African American", "Hispanic or Latino",
    "American Indian or Alaska Native", "Asian",
    "Native Hawaiian or Other Pacific Islander", "Mixed/Other/Unknown Race"
  )
)
label(demo$race) <- "Race/Ethnicity"
```

# Participant characteristics

```{r}
#| label: tbl-3
#| tbl-cap: Characteristics by race
demo %>%
  select(race) %>%
  tbl_summary()
```

@@

```{r}
#| label: fig-planets
#| fig-cap: Planets
```
